---
title: "More ggplot2: Facets, Geoms, Stats, Positioning and Coords"
author: "ST537 Group 4"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# 3.5 Facets
## Paul

```{r}

```

# 3.6 Geometric objects
## 

```{r}

```

# 3.7 Statistical transformations
## John

```{r}

```

# 3.8 Position adjustments
## 

```{r}

```

# 3.9 Coordinate systems
## Andrea

```{r}
mpg |> 
  ggplot(aes(class, hwy)) +
  geom_boxplot()

# flip x and y
mpg |> 
  ggplot(aes(class, hwy)) +
  geom_boxplot() +
  coord_flip()

# graph the world
world <- map_data("world")

world |> 
  ggplot(aes(long, lat,
             group = group)) +
  geom_polygon(fill = "steelblue", color = "black") 

world |> 
  ggplot(aes(long, lat,
             group = group)) +
  geom_polygon(fill = "steelblue", color = "black") +
  coord_quickmap()    

# pie graph
bar <- diamonds |> 
  ggplot() +
  geom_bar(aes(cut,
               fill = cut),
           show.legend = FALSE,
           width = 1) +
  theme(aspect.ratio = 1) +
  labs(x = NULL, 
       y = NULL)

bar + coord_flip()
bar + coord_polar()
```
## 3.9.1 Exercises
1.
```{r}
mpg |> 
  ggplot() +
  geom_bar(aes(x = manufacturer, 
               fill = manufacturer),
           show.legend = FALSE) +
  coord_flip() 
```

2. labs() is used to modify axis, legend, and plot labels.

3. coord_map() transforms a portion of the earth onto a flat 2D surface plane using projections within a package called mapproj. It does not preserve straight lines. coord_quickmap() works similarly, but does preserve straight lines and works best for small areas nearer to the equator.

4. The plot shows that city and highway mpg are quite positively correlated. geom_abline() adds a line that has default values of an intercept of 0 and a slope of 1. coord_fixed() imposes a specified aspect ratio onto the graph with a default option of 1, rather than automatically determined aspect ratio. 

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed(6/3)

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() +
  geom_abline() +
  coord_fixed() 
```


