---
title: "Alternative Commute Plot"
author: "ST 537 Group 4"
date: "`r Sys.Date()`"
output: pdf_document
knit: (function(inputFile, encoding) {
      out_dir <- "../Outputs";
      rmarkdown::render(inputFile,
                        encoding=encoding,
                        output_dir=file.path(dirname(inputFile), out_dir))})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r, include=FALSE}
commute <- read_rds("http://data.cwick.co.nz/commute.rds")        # import data
states <- c("or", "ak", "ca", "wa", "id", "nv")      # states of interest
commute_nw <- filter(commute, state %in% states)     # make df for graphing
```

```{r, echo=FALSE}
commute_nw |>
  filter(transport_type == "Bicycle") |>              # filter on bikes
  ggplot(aes(x = reorder(state_name, -prop),          # decending x axis
             y = prop*100,
             fill = reorder(state_name, -prop))) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("steelblue", "gray",   # make OR different color
                               "gray", "gray", "gray", "gray")) +
  theme_classic() +
  theme(legend.position = "none") +                   # hide legend
  labs(x = "State",                                   # edit graph labels
       y = "Percent of commuters",
       title = "Oregon commutes to work by bike more than other western states")
```

### Justification

We decided to only graph the bike portion of the commuter data since that was the comparison of interest in the original graph. We went with a bar chart to minimize possible reader error with a common scale to have the highest ranking on the perceptual tasks. We also used positioning and color to help with preattentive search. These are improvements over the original by using position along a common scale rather than area.
