---
title: "More ggplot2: Facets, Geoms, Stats, Positioning and Coords"
author: "ST537 Group 4"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# 3.5 Facets
## Paul

```{r}

```

# 3.6 Geometric objects
## Matt

```{r}
## 1 ##
## geom_line() , geom_boxplot() , geom_histogram() , 	geom_area() ##

## 2 ##
## displays highway mpg of vehicles with different displacements w/seperate trend lines
## of 4wd, fwd, and rwd (smoothed) ##
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)

## 3 ##
## helps hide a specific layer from being displayed. Without it that legend in the 
## layer would be visible. ##

## 4 ##
## se adds a area around the line that is shaded, representing the standard error
## with a 95% CI. ##

## 5 ##
## no these would not look different as they functionally do the same thing, just
## present the code in two differing ways. ##

## 6 ##
## Graph 1##
ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth()

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(color = "black") +
  geom_smooth(aes(group = drv), color = "blue", se = FALSE)

ggplot(mpg, aes(x = displ, y = hwy, color = drv)) +
  geom_point() +
  geom_smooth(se = FALSE)

ggplot(mpg, aes(x = displ, y = hwy, color = drv)) +
  geom_point() +
  geom_smooth(color = "blue", se = FALSE)

ggplot(mpg, aes(x = displ, y = hwy, color = drv, linetype = drv)) +
  geom_point() +
  geom_smooth(color = "blue", se = FALSE)

ggplot(mpg, aes(x = displ, y = hwy, fill = drv)) +
  geom_point(shape = 21, color = "white", size = 2) +
  geom_smooth(aes(color = drv), se = FALSE)


```

# 3.7 Statistical transformations
## John

1.	What is the default geom associated with stat_summary()? How could you rewrite the previous plot to use that geom function instead of the stat function?

The default geom associated with stat_summary() is a pointrange. You could rewrite the previous plot as follows:

```{r}
diamonds.df <- diamonds %>%
  group_by(cut) %>%
  summarize(
    median_depth = median(depth),
    min_depth = min(depth),
    max_depth = max(depth)
  )

ggplot(diamonds.df, aes(x = cut, y = median_depth, ymin = min_depth, ymax = max_depth)) +
  geom_pointrange()
```

2.	What does geom_col() do? How is it different to geom_bar()?

Geom_bar() counts the number of observations in each category, and geom_col() provides the values; it just draws bars based on the values you give it.

3.	Most geoms and stats come in pairs that are almost always used in concert. Read through the documentation and make a list of all the pairs. What do they have in common?

The common thread through these is that stat_x() transforms the data, while geom_x() handles the visualization of the data. 

4.	What variables does stat_smooth() compute? What parameters control its behavior?

Stat_smooth() computes the variables x (predictor variable), y (response variable), ymin (lower bound of confidence interval), ymax (upper bound of confidence interval), and se (whether the confidence interval is shown). The parameters that control its behavior are method (smoothing method for the line), formula (a custom model formula), se (whether to show the confidence interval), level (confidence level), span (smoothing span for method = “loess”), fullrange (whether or not to extend the line to plot limits), and geom (whether to display as smooth, line, or ribbon). 

5.	In our proportion bar chart, we need to set group = 1. Why? In other words, what is the problem with these two graphs?

Without group = 1 in the first plot, ggplot treats cut as its own group, so the proportions were calculated within each bar but not across all bars. Similarly, in the second graph, the proportions are calculated within the fill group rather than across fill groups.


# 3.8 Position adjustments
## 

```{r}

```

# 3.9 Coordinate systems
## Andrea













```{r}
mpg |> 
  ggplot(aes(class, hwy)) +
  geom_boxplot()

# flip x and y
mpg |> 
  ggplot(aes(class, hwy)) +
  geom_boxplot() +
  coord_flip()

# graph the world
world <- map_data("world")

world |> 
  ggplot(aes(long, lat,
             group = group)) +
  geom_polygon(fill = "steelblue", color = "black") 

world |> 
  ggplot(aes(long, lat,
             group = group)) +
  geom_polygon(fill = "steelblue", color = "black") +
  coord_quickmap()    

# pie graph
bar <- diamonds |> 
  ggplot() +
  geom_bar(aes(cut,
               fill = cut),
           show.legend = FALSE,
           width = 1) +
  theme(aspect.ratio = 1) +
  labs(x = NULL, 
       y = NULL)

bar + coord_flip()
bar + coord_polar()
```
## 3.9.1 Exercises
1.
```{r}
mpg |> 
  ggplot() +
  geom_bar(aes(x = manufacturer, 
               fill = manufacturer),
           show.legend = FALSE) +
  coord_flip() 
```

2. labs() is used to modify axis, legend, and plot labels.

3. coord_map() transforms a portion of the earth onto a flat 2D surface plane using projections within a package called mapproj. It does not preserve straight lines. coord_quickmap() works similarly, but does preserve straight lines and works best for small areas nearer to the equator.

4. The plot shows that city and highway mpg are quite positively correlated. geom_abline() adds a line that has default values of an intercept of 0 and a slope of 1. coord_fixed() imposes a specified aspect ratio onto the graph with a default option of 1, rather than automatically determined aspect ratio. 

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed(6/3)

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() +
  geom_abline() +
  coord_fixed() 
```


Hello!

